---
title: "Benchmarking hydrorecipes"
author: "Jonathan Kennel"
format: html
editor_options: 
  chunk_output_type: console
---

# Timings *hydrorecipes* vs *recipes*

```{r}
#| echo: false
#| warning: false
#| message: false

library(bench)
library(tibble)
library(data.table)
library(dplyr)
library(collapse)
library(recipes)
library(hydrorecipes)
library(ggbeeswarm)

```

## creating a recipe

```{r}
#| echo: true
#| warning: false

relative <- FALSE
n <- c(1e2, 1e4, 1e6)
formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows)
    bench::mark(
      frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat),
      frec2 = hydrorecipes:::recipe(formula = formula, data = dat),
      rec  = recipes::recipe(formula = formula, data = dat),
      check = FALSE
    )
  }
)

results
```

## add a step

```{r}
#| echo: true
#| warning: false

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows)
    bench::mark(
      frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCenter$new(x)),
      frec2 = hydrorecipes::recipe(formula = formula, data = dat) |>
        hydrorecipes::step_center(x),
      rec  = {recipes::recipe(formula = formula, data = dat) |>
              recipes::step_center(x)},
      check = FALSE,
      relative = relative
    )
  }
)

results
```


## step_center prep

```{r}
#| echo: true
#| warning: false

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows)
    frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepCenter$new(x))
    frec2 = hydrorecipes::recipe(formula = formula, data = dat) |>
        hydrorecipes::step_center(x)      
    rec  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_center(x)
    bench::mark(
      frec$prep(),
      frec2 |> hydrorecipes:::prep(),
      rec |> recipes::prep(),
      check = FALSE,
      min_iterations = 10,
      relative = relative
    )
  }
)

results
```

## step_center bake

```{r}
#| echo: true
#| warning: false

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows)
    frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepCenter$new(x))
    frec2  = hydrorecipes::recipe(formula = formula, data = dat) |>
      hydrorecipes::step_center(x)
    rec  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_center(x)
    bench::mark(
      frec1$prep()$bake(),
      frec2 |> hydrorecipes::prep() |> hydrorecipes::bake(),
      rec |> recipes::prep() |> recipes::bake(new_data = NULL),
      check = FALSE,
      min_iterations = 10,
      relative = relative
    )
  }
)

results
```

## step_center

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x+z)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCenter$new(x, fun = fmean))$
        plate("tbl")[["x"]],
      rec  = (recipes::recipe(formula = formula, data = dat) |>
        recipes::step_center(x) |> 
        recipes::prep() |> 
        recipes::bake(new_data = NULL))[["x"]],
      check = TRUE,
      min_iterations = 10,
      relative = relative
    )
  }
)

results
```

## step_scale

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepScale$new(x, fun = fsd, n_sd = 2L))$
        plate("tbl")[["x"]],
      rec  = (recipes::recipe(formula = formula, data = dat) |>
        recipes::step_scale(x, factor = 2L) |> 
        recipes::prep() |> 
        recipes::bake(new_data = NULL))[["x"]],
      check = TRUE,
      relative = relative,
      min_iterations = 5
    )
  }
)

results
```


## step_intercept

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)
results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = rnorm(rows))
    bench::mark(
    frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepIntercept$new())$
      plate("tbl")[["intercept"]],
    rec = (recipes::recipe(formula = formula, data = dat) |>
      recipes::step_intercept() |> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL))[["intercept"]],
      check = TRUE,
      relative = relative,
      min_iterations = 2
    )
    
  }
)

results

```

## step_normalize

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x+z)
results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = rnorm(rows),
                      z = rnorm(rows))
    
    bench::mark(
      frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepNormalize$new(c(x, y, z)))$
        plate("tbl")[, c("x","z","y")],
      
      frec2 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCenter$new(c(x, y, z)))$
        add_step(hydrorecipes:::StepScale$new(c(x, y, z)))$
        plate("tbl")[, c("x","z","y")],
      
      rec = recipes::recipe(formula = formula, data = dat) |>
        recipes::step_normalize(x, y, z) |> 
        recipes::prep() |> 
        recipes::bake(new_data = NULL),
      
      relative = relative,
      min_iterations = 2,
      check = TRUE
    )
  }
)

results


```

## step_lag

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
      frec1 = unname(hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                      add_step(hydrorecipes:::StepLeadLag$new(x, lag = 1:30))$
                      plate("tbl")),
      rec  = unname(recipes::recipe(formula = formula, data = dat) |>
                      recipes::step_lag(x, lag = 1:30) |> 
                      recipes::prep() |> 
                      recipes::bake(new_data = NULL)),
      check = TRUE,
      relative = relative,
      min_iterations = 10
    )
  }
)

results


```

## step_distributed_lag

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = c(1e5, 1e6, 1e7),
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepDistributedLag$new(x,
                                        knots = hydrorecipes:::log_lags_arma(5, 86401)))$plate(),
      # rec  = recipes::recipe(formula = formula, data = dat) |>
      #   hydrorecipes::step_distributed_lag(x, 
      #                                      knots = hydrorecipes:::log_lags_arma(5, 86401)) |> recipes::prep() |> recipes::bake(new_data = NULL),
      check = FALSE,
      relative = relative,
      min_iterations = 2
    )
  }
)

results
```

## step_harmonic

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
    frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepHarmonic$new(x, 
                                frequency = c(1.0, 2.0, 3.0), 
                                cycle_size = 0.1, 
                                starting_value = 0.0))$
      plate("tbl"),
    rec  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_harmonic(x, 
                             frequency = c(1.0, 2.0, 3.0), 
                             cycle_size = 0.1, 
                             starting_val = 0.0,
                             keep_original_cols = TRUE) |> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL),
    
    # sin and cos terms order is different
      check = FALSE,
      relative = relative,
      min_iterations = 10
    )
  }
)
results
```

## step_pca

```{r}
#| echo: true
#| warning: false
set.seed(1)
formula <- as.formula(x~a+b+c+d+e+f+g+h+i+j+k+l)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      a = rnorm(rows),
                      b = rnorm(rows),
                      c = rnorm(rows),
                      d = rnorm(rows),
                      e = rnorm(rows),
                      f = rnorm(rows),
                      g = rnorm(rows),
                      h = rnorm(rows),
                      i = rnorm(rows),
                      j = rnorm(rows),
                      k = rnorm(rows),
                      l = rnorm(rows)
    )
    bench::mark(
    frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPca$new(all_predictor(),
                           n_comp = 10L))$
      plate(),
    frec2 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPca$new(all_predictor(),
                           n_comp = 5L))$
      plate(),
    frec3 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPca$new(all_predictor(),
                           n_comp = 10L,
                           center = FALSE,
                           scale = FALSE))$
      plate(),
    frec4 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPca$new(all_predictor(),
                           n_comp = 5L,
                           center = FALSE,
                           scale = FALSE))$
      plate(),
    rec1  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_pca(recipes::all_predictors(),
                        num_comp = 10L,
                        options = list(center = TRUE, scale. = TRUE))|> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL),
    rec2  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_pca(recipes::all_predictors(),
                        num_comp = 5L,
                        options = list(center = TRUE, scale. = TRUE)) |> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL),
    rec3  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_pca(recipes::all_predictors(),
                        num_comp = 10L) |> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL),
    rec4  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_pca(recipes::all_predictors(),
                        num_comp = 5L) |> 
      recipes::prep() |> 
      recipes::bake(new_data = NULL),
      check = FALSE,
      relative = relative,
      min_iterations = 2
    )
  }
)


print(results, n = 100)



```


## step_dummy

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = qF(sample(1:10, rows, replace = TRUE)),
                      z = rnorm(rows))
    bench::mark(
      frec = unname(hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepDummy$new(y))$plate("tbl")),
      rec  = unname(recipes::recipe(formula = formula, data = dat) |>
        recipes::step_dummy(y, keep_original_cols = TRUE) |>
        recipes::prep() |>
        recipes::bake(new_data = NULL)),
      check = TRUE,
      relative = FALSE,
      min_iterations = 1
    )
  }
)

results


```

## step_find_interval

- no direct comparison so compare to step_cut

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepFindInterval$new(x, vec = c(-0.1, 0, 0.1)))$
        plate("tbl"),
      rec  = recipes::recipe(formula = formula, data = dat) |>
        recipes::step_cut(x, breaks = c(-0.1, 0, 0.1)) |>
        recipes::prep() |>
        recipes::bake(new_data = NULL),
      check = FALSE,
      relative = relative,
      min_iterations = 1
    )
  }
)

results


```



## step_varying

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x+z)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rep(1, rows), 
                      y = 1:rows,
                      z = rnorm(rows))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepVarying$new(c(x, y, z)))$
        plate("tbl"),
      rec  = recipes::recipe(formula = formula, data = dat) |>
        recipes::step_zv(x, y, z) |>
        recipes::prep() |>
        recipes::bake(new_data = NULL),
      check = TRUE,
      relative = relative,
      min_iterations = 1
    )
  }
)

results


```

## step_kernel_filter

```{r}
#| echo: true
#| warning: false

library(RcppRoll)
formula <- as.formula(y~x+z)

results <- bench::press(
  rows = c(1e4, 1e5),
  {
    dat <- data.frame(x = rep(1, rows), 
                      y = 1:rows,
                      z = cumsum(rnorm(rows)))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepKernelFilter$new(z, 
                                      kernel = list(rep(1, 5001)/5001), align = "center"))$
        plate("tbl"),
      rec  = recipes::recipe(formula = formula, data = dat) |>
        recipes::step_window(z, size = 5001, statistic = "mean") |>
        recipes::prep() |>
        recipes::bake(new_data = NULL),
      min_iterations = 1,
      check = FALSE
    )
  }
)

results


```

## multiple steps

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = as.numeric(1:rows), 
                      y = 1:rows)
    frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepLeadLag$new(x, lag = 1:10))$
      add_step(hydrorecipes:::StepHarmonic$new(x, 
                                frequency = c(1,2,3,4,5), 
                                cycle_size = 0.1, 
                                starting_value = 0))$
      add_step(hydrorecipes:::StepCenter$new(x))
    rec  = recipes::recipe(formula = formula, data = dat) |>
      recipes::step_lag(x, lag = 1:10, keep_original_cols = TRUE) |>
      recipes::step_harmonic(x, 
                             frequency = c(1,2,3,4,5), 
                             cycle_size = 0.1, 
                             starting_val = 0,
                             keep_original_cols = TRUE) |>
      recipes::step_center(x)
    bench::mark(
      frec$plate("tbl"),
      rec |> recipes::prep() |> recipes::bake(new_data = NULL),
      check = FALSE,
      relative = relative,
      min_iterations = 1
    )
  }
)

results


```


## step_spline_b

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows),
                      y = 1:rows)
    bench::mark(
      frec = unname(hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                      add_step(hydrorecipes:::StepSplineB$new(x, df = 13))$
                      plate("tbl")),
      rec  = unname(recipes::recipe(formula = formula, data = dat) |>
                      recipes::step_spline_b(x, deg_free = 13, keep_original_cols = TRUE)|> 
                      recipes::prep() |> 
                      recipes::bake(new_data = NULL)),
      check = TRUE,
      relative = relative,
      min_iterations = 10
    )
  }
)

results


start <- as.POSIXct("2016-08-18", tz = "UTC")
end <- as.POSIXct("2016-10-13", tz = "UTC")

dates <- data.table(datetime = seq.POSIXt(start, end, 1))

aa <- recipe(datetime~., dates) |> 
  step_spline_b(datetime, df = 13) |>
  prep() |> 
  bake()

```

## step_aquifer_grf & step_aquifer_theis

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = as.numeric(1:rows),
                      y = rep(0.01, rows))
    bench::mark(
      frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferGRF$new(time = x,
                                                        flow_rate = y))$
                 plate("dt")),
      frec2 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferTheis$new(time = x,
                                                          flow_rate = y))$
                 plate("dt")),
      check = TRUE,
      relative = relative)
  }
)

results


```

## step_aquifer_leaky

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = as.numeric(1:rows),
                      y = rep(0.01, rows))
    bench::mark(
      frec1 = unname(hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferLeaky$new(time = x,
                                                          flow_rate = y,
                                                          leakage = 100000000))$
                 plate("dt")),
      frec2 = unname(hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferTheis$new(time = x,
                                                          flow_rate = y))$
                 plate("dt")),
      check = TRUE,
      relative = relative)
  }
)

results


```

## step_aquifer_patch

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = c(1e5),
  {
    dat <- data.frame(x = as.numeric(1:rows),
                      y = rep(0.01, rows))
    bench::mark(
      frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferGRF$new(time = x,
                                                        flow_rate = y))$
                 plate("dt")),
      frec3 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepAquiferPatch$new(time = x,
                                                          flow_rate = 0.01,
                                                          thickness = 1.0,
                                                          radius = 100.0,
                                                          radius_patch = 200.0,
                                                          specific_storage_inner = 1e-6,
                                                          specific_storage_outer = 1e-6,
                                                          hydraulic_conductivity_inner = 1e-4,
                                                          hydraulic_conductivity_outer = 1e-4,
                                                          n_stehfest = 8L
                 ))$
                 plate("dt")),
      check = FALSE,
      relative = relative)
  }
)


results
```

## step_vadose_weeks

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = as.numeric(1:rows), 
                      y = as.numeric(1:rows))
    bench::mark(
      frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepVadoseWeeks$new(time = x, 
                                              air_diffusivity = 0.8, 
                                              thickness = 5, 
                                              precision = 1e-12))$
                 plate("dt")),
      check = FALSE,
      min_iterations = 10
    )
  }
)

results
```

## step_transport_ogata_banks

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(expand.grid(as.numeric(1:rows), as.numeric(1:10)))
    names(dat) <- c('x', 'y')
    bench::mark(
      frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
                 add_step(hydrorecipes:::StepTransportOgataBanks$new(time = x,
                                                                 distance = y))$
                 plate("dt")),
      check = FALSE,
      min_iterations = 10
    )
  }
)

results
```

## step_transport_fractures_solute
```{r}
#| echo: true
#| warning: false

formula <- as.formula(~time+z+x)

dat <- setDT(expand.grid(10^(3:8),
                         seq(0.0, 10, 1),
                         c(0.0)))

names(dat) <- c("time", "z", "x")

results <- 
  bench::mark(
    frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
               add_step(hydrorecipes:::StepTransportFracturesSolute$new(time = time,
                                                         distance_fracture = z,
                                                         distance_matrix = x))$
               plate("dt")),
    check = FALSE,
    min_iterations = 10
  )

results
```

## step_transport_fractures_heat

```{r}
#| echo: true
#| warning: false

formula <- as.formula(~time+z+x)

dat <- setDT(expand.grid(10^(3:8),
                         seq(0.0, 100, 1),
                         c(0.0, 0.05)))

names(dat) <- c("time", "z", "x")

results <- 
  bench::mark(
    frec1 = (hydrorecipes:::Recipe$new(formula = formula, data = dat)$
               add_step(hydrorecipes:::StepTransportFracturesHeat$new(time = time,
                                                         distance_fracture = z,
                                                         distance_matrix = x))$
               plate("dt")),
    check = FALSE,
    min_iterations = 10
  )

results
```

## step_pgram and step_welch

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x + z)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), y = rnorm(rows), z = rnorm(rows),
                      q = rnorm(rows), r = rnorm(rows), s = rnorm(rows))
    frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPgram$new(c(x, y), 
                             3,
                             TRUE,
                             TRUE,
                             FALSE,
                             0.1))
    frec3 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepPgram$new(c(x, y), 
                             3,
                             TRUE,
                             TRUE,
                             TRUE,
                             0.1))
    frec2 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepWelch$new(c(x, y),
                             length_subset =  nrow(dat) / 10,
                             overlap = 0.60,
                             window = window_nuttall(nrow(dat) / 10)))
    bench::mark(
      frec3$plate("dt"),
      frec1$plate("dt"),
      frec2$plate("dt"),
      check = FALSE,
      min_iterations = 3
    )
  }
)

results
```

## step_transfer_welch and step_transfer_pgram

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), y = rnorm(rows))
    frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepTransferPgram$new(c(x, y), 
                                     3,
                                     TRUE,
                                     TRUE,
                                     0.1))
    frec2 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
      add_step(hydrorecipes:::StepTransferWelch$new(c(x, y),
                                     length_subset =  nrow(dat) / 10,
                                     overlap = 0.60,
                                     window = window_nuttall(nrow(dat) / 10)))
    bench::mark(frec1$plate("dt"),
                frec2$plate("dt"),
                check = FALSE,
                min_iterations = 2
    )
  }
)

results
```


## step_ols

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x+z)


results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows), 
                      y = rnorm(rows),
                      z = rnorm(rows))
    bench::mark(
      frec = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCenter$new(x))$
        add_step(hydrorecipes:::StepScale$new(z))$
        add_step(hydrorecipes:::StepOls$new(formula = formula))$
        plate(),
      check = FALSE,
      relative = relative
    )
  }
)

results
```


## step_ols_gap_fill

```{r}
#| echo: true
#| warning: false

set.seed(123)
n <- 100000
frm <- formula(x ~ y + z)
x <- cumsum(rnorm(n))
dat <- data.table(x = x, y = x, z = as.numeric(1:n))
dat[, x := x + c(rep(20, n/2), rep(0, n/2))]
dat[, x := x + 3.0 * sin(z * 1/n)]
tmp <- copy(dat$x)
plot(tmp, type = 'l')
dat[60000:70000, x := NA_real_]

dat <- unclass(dat)
f = hydrorecipes:::Recipe$new(formula = frm, data = (dat))$
  add_step(hydrorecipes:::StepFindInterval$new(z, vec = c(0, n/2, n)))$
  add_step(hydrorecipes:::StepIntercept$new())$
  add_step(hydrorecipes:::StepSplineB$new(z, df = 4))$
  add_step(hydrorecipes:::StepDropColumns$new(z))

frec = hydrorecipes:::Recipe$new(formula = frm, data = dat)$
  add_step(hydrorecipes:::StepOlsGapFill$new(c(x, y, z), recipe = f))


tmp2 <- frec$plate()
points(dat$x, type = 'p', pch = 20, cex = 0.2, col = "yellow")
points(tmp2$ols_gap_fill_x, type = 'l', col = 'red')





```



# check

## step_check_spacing

```{r}
#| echo: true
#| warning: false

formula <- as.formula(y~x)

results <- bench::press(
  rows = n,
  {
    dat <- data.frame(x = rnorm(rows),
                      y = 1:rows)
    dat[9, "x"] <- NA
    dat[9, "y"] <- NA
    
    bench::mark(
      frec1 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCheckSpacing$new(y))$
        add_step(hydrorecipes:::StepCheckNA$new(y))$
        prep()$
        bake(),
      frec2 = hydrorecipes:::Recipe$new(formula = formula, data = dat)$
        add_step(hydrorecipes:::StepCheckSpacing$new(x))$
        add_step(hydrorecipes:::StepCheckNA$new(x))$
        prep()$
        bake(),
      check = FALSE,
      relative = relative,
      min_iterations = 10
    )
  }
)

results



```


<!-- ```{r} -->

<!-- x <- alloc(1.0, 1e6) # Vector -->

<!-- bench::mark( x %*=% 1.0, x %/=% 1.0,x %-=% 0.0, x %+=% 0.0, 1.0/1.0, check = FALSE) -->


<!-- n <- 1000000 -->
<!-- input <- data.frame(i = 1:n, -->
<!--                     c = sample(letters, n, replace = TRUE), -->
<!--                     f = factor(sample(letters, n, replace = TRUE)), -->
<!--                     d = rnorm(n)) -->
<!-- bench::mark(input[[1]], -->
<!--             unclass(input)[[1]], -->
<!--             get_vars(input, 1),  -->
<!--             get_elem(input, 1, recursive = FALSE), -->
<!--             check = FALSE, relative = TRUE) -->

<!-- initial <- data.frame(z = 1:n) -->
<!-- tmp <- findInterval(input[['f']],c(3,7)) -->
<!-- lst <- as.list(input)  -->
<!-- m   <- as.matrix(input) -->
<!-- dt  <- as.data.table(input) -->
<!-- tbl <- as_tibble(input) -->
<!-- df  <- input -->
<!-- mm  <- qDF(m) -->

<!-- # m <- matrix(rnorm(3e6), ncol = 3) -->
<!-- # df <- qDF(m) -->
<!-- # bench::mark(colMeans(m),as.numeric(colMeans(df)), fmean(m), as.numeric(fmean(df)), check = FALSE) -->

<!-- microbenchmark( -->
<!--   bind_cols(initial, lst, .name_repair = 'minimal'), -->
<!--   bind_cols(initial, dt, .name_repair = 'minimal'), -->
<!--   bind_cols(initial, unclass(dt), .name_repair = 'minimal'), -->
<!--   bind_cols(initial, m, .name_repair = 'minimal'), -->
<!--   bind_cols(initial, tbl, .name_repair = 'minimal'), -->
<!--   bind_cols(initial, df, .name_repair = 'minimal'), -->

<!--   cbind(initial, qDF(lst), stringsAsFactors = FALSE), -->
<!--   cbind(initial, qDF(dt), stringsAsFactors = FALSE), -->
<!--   cbind(initial, qDF(unclass(dt)), stringsAsFactors = FALSE), -->
<!--   cbind(initial, qDF(m), stringsAsFactors = FALSE), -->
<!--   cbind(initial, qDF(tbl), stringsAsFactors = FALSE), -->
<!--   cbind(initial, qDF(df), stringsAsFactors = FALSE), -->

<!--   cbind(initial, (lst), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (dt), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (unclass(dt)), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (m), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (tbl), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (df), stringsAsFactors = FALSE), -->


<!--   initial[['a']] <- m[,1], -->
<!--   cbind(cbind(initial, df), df), -->
<!--   cbind(cbind(initial, qDF(lst)), qDF(lst)), -->
<!--   cbind(cbind(initial, tbl), tbl), -->
<!--   bind_cols(bind_cols(initial, lst), lst, .name_repair = 'minimal'), -->
<!--   times = 3 -->
<!-- ) -->

<!-- bench::mark(   -->
<!--   to_dummy_lst(tmp, 3), -->
<!--   check = FALSE -->
<!-- ) -->

<!-- n <- 5 -->
<!-- microbenchmark::microbenchmark( -->
<!--   to_dummy(tmp, n), -->
<!--   to_dummy2(tmp, n), -->
<!--   to_dummy3(tmp, n), -->
<!--   to_dummy4(tmp, n), -->
<!--   to_dummy5(tmp, n), -->
<!--   to_dummy_list(tmp, n), -->
<!--   # to_dummy6(tmp, n), -->
<!--   # to_dummy_lst(tmp, n, paste0('x', 1:n)), -->
<!--   # cbind(initial,to_dummy_lst2(cbind(tmp, tmp), n, paste0('x', 1:n))), -->
<!--   # to_dummy_df(tmp, n, paste0('x', 1:n)), -->
<!--   times = 3 -->
<!-- ) -->

<!-- bench::mark( -->
<!--   qDF(to_dummy(tmp, 10)), -->
<!--   to_dummy_lst(tmp, 10, paste0('x', 1:10)), -->
<!--   to_dummy_df(tmp, 10, paste0('x', 1:10)), -->
<!--   check = FALSE -->
<!-- ) -->
<!-- n <- 10 -->
<!-- microbenchmark::microbenchmark( -->
<!--   cbind(initial, qDF(to_dummy(tmp, n))), -->
<!--   cbind(initial, qDF(to_dummy3(tmp, n))), -->
<!--   cbind(initial, to_dummy_lst(tmp, n, paste0('x', 1:n))), -->
<!--   cbind(initial, to_dummy_df(tmp, n, paste0('x', 1:n))), -->
<!--   times = 5 -->
<!-- ) -->

<!-- bench::mark( -->
<!--   cbind(initial, qDF(to_dummy(tmp, 6))), -->
<!--   cbind(initial, to_dummy_lst(tmp, 6, paste0('x', 1:6))), -->
<!--   cbind(initial, to_dummy_df(tmp, 6, paste0('x', 1:6))), -->
<!--   check = FALSE -->
<!-- ) -->

<!-- bench::mark( -->
<!--   cbind(initial, (lst), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (df), stringsAsFactors = FALSE), -->
<!--   cbind(initial, (tbl), stringsAsFactors = FALSE), -->
<!--   check = FALSE -->
<!-- ) -->

<!-- str(cbind(initial, (lst), stringsAsFactors = FALSE)) -->
<!-- str(cbind(initial, (df), stringsAsFactors = FALSE)) -->


<!-- n <- 1000000L -->
<!-- m <- as.matrix(1:n) -->
<!-- d <- sample(1:10, 1e5, replace = TRUE) -->

<!-- v <- 1:n -->
<!-- nn <- 10 -->
<!-- microbenchmark::microbenchmark( -->
<!--   m0 <- lag_matrix(m, 1:nn, suffix = 'a', prefix = 'b', n_subset = 1), -->
<!--   m1 <- lag_matrix(m, 1:nn, suffix = 'a', prefix = 'b', n_subset = 5), -->
<!--   m3 <- lag_lead_matrix(v, 1:nn, 1, 0), -->
<!--   m4 <- lag_lead_matrix(v, 1:nn, 5, 0), -->
<!--   times = 3 -->
<!-- ) -->

<!-- bench::mark( -->
<!--   m0 <- lag_matrix(m, 1:nn, suffix = 'a', prefix = 'b', n_subset = 1), -->
<!--   m1 <- lag_matrix(m, 1:nn, suffix = 'a', prefix = 'b', n_subset = 5), -->
<!--   m3 <- lag_lead_matrix(v, 1:nn, 1, 0), -->
<!--   m4 <- lag_lead_matrix(v, 1:nn, 5, 0), -->
<!--   check = FALSE -->
<!-- ) -->



<!-- n <- 10000L -->
<!-- t <- sort(rnorm(n)) -->
<!-- m <- matrix(1:(8*n), ncol = 8) -->
<!-- bench::mark( -->
<!--   h1 <- harmonic(t, c(1,2,4,5)), -->
<!--   h2 <- harmonic_float(t, c(1,2,4,5)), -->
<!--   h3 <- harmonic_field(t, c(1,2,4,5)), -->
<!--   check = FALSE -->
<!-- ) -->

<!-- microbenchmark::microbenchmark( -->
<!--   b <- harmonic(t, c(1,2,4,5, 8, 9)), -->
<!--   a <- harmonic_float(t, c(1,2,4,5, 8, 9)), -->
<!--   c <- harmonic_field(t, c(1,2,4,5, 8, 9)), -->
<!--   times = 3 -->
<!-- ) -->
<!-- ``` -->
