---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hydrorecipes

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/hydrorecipes)](https://CRAN.R-project.org/package=hydrorecipes)
[![CRAN download count](http://cranlogs.r-pkg.org/badges/grand-total/hydrorecipes?color=green)](https://cran.r-project.org/package=hydrorecipes)
[![R-CMD-check](https://github.com/jkennel/hydrorecipes/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jkennel/hydrorecipes/actions/workflows/R-CMD-check.yaml)
[![DOI](https://zenodo.org/badge/436375294.svg)](https://zenodo.org/badge/latestdoi/436375294)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Codecov test coverage](https://codecov.io/gh/jkennel/hydrorecipes/branch/master/graph/badge.svg)](https://app.codecov.io/gh/jkennel/hydrorecipes?branch=master)


<!-- badges: end -->

The goal of **hydrorecipes** is to supplement the [recipes package](https://recipes.tidymodels.org) with a few steps
that can help deal with moderately sized water level datasets. These were developed primarily with regression deconvolution in mind and used *lm* or *glmnet*, but other model engines could also be used. The following steps are currently available:

- *step_lead_lag* is a more flexible version of *step_lag* from the [recipes package](https://recipes.tidymodels.org).  Values can be negative which indicates the vector is leading.  In addition, subsetting can be done on the dataset during this process when dealing with very large datasets leading to performance gains without sacrificing temporal accuracy. 
- *step_distributed_lag* is a distributed lag approach for modelling the response in a flexible yet concise manner.  This is useful when you have a long maximum lag or large datasets.
- *step_earthtide* uses the [earthtide package](https://CRAN.R-project.org/package=earthtide) to model the synthetic Earth tide given locations and times. This can provide a single Earth tide curve or a set of harmonics that can be used in regression models.

## Installation

You can install the development version of **hydrorecipes** from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jkennel/hydrorecipes")
```

## Example

This is the method of [Kennel 2020](http://hdl.handle.net/10214/17890) which uses a distributed lag model and the [earthtide package](https://CRAN.R-project.org/package=earthtide) to generate synthetic wave groups. A ~1.5 month dataset of water and barometric pressure having a monitoring frequency of 2 minutes is presented below. The barometric response is modeled over two days using a distributed lag model with 15 regressor terms. The *knots* are logarithmically separated over two days to accurately capture early and late time responses which can be caused by different physical mechanisms.

```{r message = FALSE}
library(hydrorecipes)
library(earthtide)
library(tidyr)
library(ggplot2)

data(transducer)

# convert to numeric because step_ns doesn't handle POSIXct
transducer$datetime_num <- as.numeric(transducer$datetime)

unique(diff(transducer$datetime_num)) # times are regularly spaced

# Earth tide inputs
wave_groups <- earthtide::eterna_wavegroups
wave_groups <- na.omit(wave_groups[wave_groups$time == '1 month', ])
wave_groups <- wave_groups[wave_groups$start > 0.5, ]
latitude    <-   34.0
longitude   <- -118.5

# create recipe 
rec <- recipe(wl~baro+datetime_num, transducer) |>
  step_distributed_lag(baro, knots = log_lags(15, 86400 * 2 / 120)) |>
  step_earthtide(datetime_num,
                 latitude = latitude,
                 longitude = longitude,
                 astro_update = 1,
                 wave_groups = wave_groups) |>
  step_ns(datetime_num, deg_free = 10) |>
  prep()

input <- rec |> bake(new_data = NULL)
```

```{r}
summary(fit <- lm(wl~., input))
```


## Decomposition

The decomposition consists of:

- **Intercept** A single offset value (dbar)
- **Background Trend** The background trend determined using natural splines (dbar)
- **Barometric Component** The barometric component of the water pressure determined using distributed lag terms (dbar) 
- **Earth tide Component** The Earth tides component determined using harmonic analysis (dbar)
- **Predicted** The sum of all the components above.  This is the predicted water pressure. (dbar)
- **Water Pressure** The observed water pressure measured with a non-vented transducer (dbar)
- **Residuals (obs-mod)** The difference between the observed water pressure and the modeled pressure (Predicted). (dbar)

```{r fig.height = 11, fig.width = 12, out.width = '90%', warning = FALSE}
pred <- predict_terms(fit = fit, 
                      rec = rec,
                      data = input)
pred <- bind_cols(transducer[, c('datetime', 'wl')], pred)
pred$residuals <- pred$wl - pred$predicted
pred_long <- pivot_longer(pred, cols = !datetime)
levels <-c('intercept', 'ns_datetime_num', 'distributed_lag_baro',
           'earthtide_datetime_num', 'predicted', 'wl', 'residuals')
labels <- c('Intercept', 'Background trend', 'Barometric Component',
            'Earth Tide Component', 'Predicted', 'Water Pressure', 
            'Residuals (obs-mod)')
pred_long$name <- factor(pred_long$name, 
                         levels = levels,
                         labels = labels)
ggplot(pred_long, aes(x = datetime, y = value)) +
  geom_line() + 
  scale_y_continuous(labels = scales::comma) + 
  scale_x_datetime(expand = c(0,0)) + 
  ggtitle('Water Level Decomposition Results') + 
  xlab("") + 
  facet_grid(name~., scales = 'free_y') + 
  theme_bw()
```

## Response

There are two responses for this model:

- **distributed_lag** The barometric loading response.  This is an impulse response function and the cumulative response is usually presented.  This value typically ranges between 0 and 1 and represents the expected water pressure/level change following a 1 unit change in barometric pressure.  The same units for water pressure and barometric pressure should be used (Figure A).
- **earthtide** The Earth tide response is represented by a set of harmonic wave groups.  These results are usually presented as amplitude and phase as a function of frequency.  Depending on the length of your dataset you may want to include more harmonic wavegroups.  For locations with minimal ocean tide effects it may be useful to compare the relative amplitudes to theory as an assessment of the model quality (how well all other signals were removed). The phase shift for small amplitude components may not be reliable (Figure B).  


```{r fig.height = 5, fig.width = 5, out.width = '47%', fig.show='hold'}
resp    <- response(fit, rec)
resp_ba <- resp[resp$name == 'cumulative', ]
resp_ba <- resp_ba[resp_ba$term == 'baro', ]
ggplot(resp_ba, aes(x = x * 120 / 3600, y = value)) +
  ggtitle('A: Barometric Loading Response') + 
  xlab('lag (hours)') +
  ylab('Cumulative Response') +
  scale_y_continuous(limits = c(0, 1)) +
  geom_line() + 
  theme_bw()
resp_et <- resp[resp$name %in% c('amplitude', 'phase'), ]
ggplot(resp_et, aes(x = x, xend = x, y = 0, yend = value)) +
  geom_segment() + 
  ggtitle('B: Earthtide Response') +
  xlab('Frequency (cycles per day)') +
  ylab('Phase (radians)   |   Amplitude (dbar)') +
  facet_grid(name~., scales = 'free_y') + 
  theme_bw()
```
