---
title: "Barometric pressure steps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Barometric pressure steps}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results = "hide", include = FALSE}
library(hydrorecipes)
```

```{r data}
# minute data
data(kennel_2020)
```

### Time domain

#### Single value

```{r clark}
clarks <- recipe(wl~., kennel_2020) |>
  step_baro_clark(wl,
                  baro,
                  lag_space = 1) |> # 1 minutes (every minute differences)
  step_baro_clark(wl,
                  baro, 
                  lag_space = 60) |> # 60 minutes (hourly differences)
  step_baro_clark(wl, 
                  baro, 
                  lag_space = 1440) |> # 1440 minutes (daily differences)
  prep() |>
  bake()

# The barometric efficiency is stored in the step (warning: this may change)
clarks$get_step_data("barometric_efficiency")

```

```{r regression}
least_squares <- recipe(wl~., kennel_2020) |>
  step_baro_least_squares(wl, baro) |> 
  step_baro_least_squares(wl, baro, 
                          lag_space = 1440, 
                          differences = TRUE) |> # 1440 minutes (daily differences)
  prep() |>
  bake()

least_squares$get_step_data("barometric_efficiency")


```

#### Response function

Standard lag method

```{r lag}

lag_standard <- recipe(wl~., kennel_2020) |>
  step_lead_lag(baro, lag = 0:1440) |> 
  step_spline_b(datetime, df = 11, intercept = TRUE) |> 
  step_drop_columns(c(baro, datetime, et)) |>
  step_ols(wl~.) |> 
  prep() |>
  bake()
resp_l <- lag_standard$get_response_data()

```

Distributed lag method

```{r distlag, fig.alt = "Barometric response functions with standard leading and lagging method compared to the distributed lag method.  Both methods are generally the same but the distributed lag method is smoother."}
lag_standard <- recipe(wl~., kennel_2020) |>
  step_distributed_lag(baro, knots = log_lags_arma(15, 1440)) |> 
  step_spline_b(datetime, df = 11, intercept = TRUE) |> 
  step_drop_columns(c(baro, datetime, et)) |>
  step_ols(wl~.) |> 
  prep() |>
  bake()
resp_dl <- lag_standard$get_response_data()

plot(value~x, resp_l[resp_l$term == "lead_lag" & resp_l$variable == "cumulative", ],
     type = "l", ylim = c(0, 1),
     xlab = "Lag in minutes",
     ylab = "Cumulative Response"
     )
points(value~x, resp_dl[resp_dl$term == "distributed_lag_interpolated" & resp_dl$variable == "cumulative", ],
       type = "l", col = "red")

```

### Frequency domain

#### Single value

These methods are based on the 2 cpd frequency which may not be indicative of the static barometric efficiency, in cases with a frequency dependent response (large wellbore storage or non-confined conditions).

```{r ac}
data(kennel_2020)

harmonic <- recipe(wl~., kennel_2020) |>
  step_baro_harmonic(datetime, wl, baro, et) |> 
  prep() |>
  bake()

harmonic$get_step_data("barometric_efficiency")

```


#### Transfer function

Three methods:

- pgram based
  - assumes smooth periodogram
- Welch's 
  - loss of low frequency information
- Experimental 
  - assumes smooth periodogram
  - smoothing across multiple frequencies
  - many higher frequecies discarded
  - should be very fast

```{r tf, fig.alt = "Transfer functions."}

data(kennel_2020)
formula <- as.formula(datetime~baro+wl+et)
formula_1 <- as.formula(wl~baro + et)

pg = recipe(formula = formula, data = kennel_2020) |>
  step_fft_transfer_welch(formula = formula_1, 
                          length_subset = 40000,
                          window = hydrorecipes:::window_rectangle(40000),
                          time_step = 60) |>
  step_fft_transfer_pgram(formula = formula_1, 
                          spans = c(3, 3), 
                          time_step = 60) |>
  step_fft_transfer_experimental(formula = formula_1, 
                                 n_groups = 600,
                                 spans = c(3, 3), 
                                 time_step = 60) |>
  prep("df") |>
  bake()

pg$get_elapsed_times()
tf <- pg$get_transfer_data(type = "dt")

plot(Mod(value)~frequency, 
     data = tf[grepl("fft_transfer_experimental", id) & variable == "wl_baro_et_1"], 
     type = "l", 
     log = "x", 
     xlim = c(0.01, 200),
     ylim = c(0, 1),
     col = "#00900080")
points(Mod(value)~frequency, 
       data = tf[grepl("fft_transfer_welch", id) & variable == "wl_baro_et_1"], 
       type = "l",
       col = "#90000050")
points(Mod(value)~frequency, 
       tf[grepl("fft_transfer_pgram", id) & variable == "wl_baro_et_1"],
       type = "l", 
       col = "#00009050")




plot(Mod(value)~frequency, 
     data = tf[grepl("fft_transfer_experimental", id) & variable == "wl_baro_et_1"], 
     type = "l", 
     log = "x", 
     xlim = c(0.01, 200),
     ylim = c(0, 1),
     col = "#00900080")

plot(Arg(value)~frequency, 
     data = tf[grepl("fft_transfer_experimental", id) & variable == "wl_baro_et_1"], 
     type = "l", 
     log = "x", 
     xlim = c(0.01, 200),
     ylim = c(0, 1),
     col = "#00900080")

```


### References


```{r session}
sessionInfo()
```
