---
title: "Barometric pressure steps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Barometric pressure steps}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results = "hide", include = FALSE}
library(hydrorecipes)
```

```{r data}
# minute data
data(kennel_2020)
```

### Time domain

#### Single value

```{r clark}
clarks <- recipe(wl~., kennel_2020) |>
  step_baro_clark(wl,
                  baro,
                  lag_space = 1) |> # 1 minutes (every minute differences)
  step_baro_clark(wl,
                  baro, 
                  lag_space = 60) |> # 60 minutes (hourly differences)
  step_baro_clark(wl, 
                  baro, 
                  lag_space = 1440) |> # 1440 minutes (daily differences)
  prep() |>
  bake()

# The barometric efficiency is stored in the step (warning: this may change)
clarks$get_step_data("barometric_efficiency")

```

```{r regression}
least_squares <- recipe(wl~., kennel_2020) |>
  step_baro_least_squares(wl, baro) |> 
  step_baro_least_squares(wl, baro, 
                          lag_space = 1440, 
                          differences = TRUE) |> # 1440 minutes (daily differences)
  prep() |>
  bake()

least_squares$get_step_data("barometric_efficiency")


```

#### Response function

Standard lag method

```{r lag}

lag_standard <- recipe(wl~., kennel_2020) |>
  step_lead_lag(baro, lag = 0:1440) |> 
  step_spline_b(datetime, df = 11, intercept = TRUE) |> 
  step_drop_columns(c(baro, datetime, et)) |>
  step_ols(wl~.) |> 
  prep() |>
  bake()
resp_l <- lag_standard$get_response_data()

```

Distributed lag method

```{r distlag}
lag_standard <- recipe(wl~., kennel_2020) |>
  step_distributed_lag(baro, knots = log_lags_arma(15, 1440)) |> 
  step_spline_b(datetime, df = 11, intercept = TRUE) |> 
  step_drop_columns(c(baro, datetime, et)) |>
  step_ols(wl~.) |> 
  prep() |>
  bake()
resp_dl <- lag_standard$get_response_data()

plot(value~x, resp_l[resp_l$term == "lead_lag" & resp_l$variable == "cumulative", ],
     type = "l", ylim = c(0, 1),
     xlab = "Lag in minutes",
     ylab = "Cumulative Response"
     )
points(value~x, resp_dl[resp_dl$term == "distributed_lag_interpolated" & resp_dl$variable == "cumulative", ],
       type = "l", col = "red")

```

### Frequency domain

#### Single value

These methods are based on the 2 cpd frequency which may not be indicative of the static barometric efficiency, in cases with a frequency dependent response (large wellbore storage or non-confined conditions).

```{r ac}
data(kennel_2020)

harmonic <- recipe(wl~., kennel_2020) |>
  step_baro_harmonic(datetime, wl, baro, et) |> 
  prep() |>
  bake()

harmonic$get_step_data("barometric_efficiency")

```


#### Transfer function

Three methods:

- pgram based
  - assumes smooth periodogram
- Welch's 
  - loss of low frequency information
- Experimental 
  - assumes smooth periodogram
  - smoothing across multiple frequencies
  - many higher frequecies discarded
  - should be very fast

```{r tf}

form <- as.formula("wl~.")

pgram <- recipe(form, kennel_2020) |>
  step_fft_transfer_pgram(c(wl, baro, et), spans = c(3, 5, 7), taper = 0.4,
                          time_step = 60) |>
  prep() |>
  bake()

pg <- as.data.frame(pgram$get_step_data("fft_result")[[1]])



welch <- recipe(form, kennel_2020) |>
  step_fft_transfer_welch(c(wl, baro, et),
                          length_subset = 40000,
                          overlap = 0.97,
                          window = window_nuttall(40000), 
                          time_step = 60) |>
  prep() |>
  bake()

we <- as.data.frame(welch$get_step_data("fft_result")[[1]])

form <- as.formula("wl~.")

pgram_experimental <- recipe(form, kennel_2020) |>
  step_fft_transfer_experimental(c(wl, baro, et), spans = c(3, 5, 7), 
                                 taper = 0.4, power = 3, n_groups = 400,
                          time_step = 60) |>
  prep() |>
  bake()

experimental <- as.data.frame(pgram_experimental$get_step_data("fft_result")[[1]])


plot(Mod(fft_transfer_pgram_1)~frequency, pg, type = "l", log = "x",
     col = "#00900080")
points(Mod(transfer_welch_1)~frequency, we, type = "l",
       col = "#90000080")
points(Mod(fft_transfer_experimental_1)~frequency, experimental, type = "l", 
       col = "#00009080")

```


### References


```{r session}
sessionInfo()
```
