---
title: "Generalized radial flow model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generalized radial flow model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results = "hide", include = FALSE}

library(hydrorecipes)
library(plotly)
set.seed(1)
```

## Solution

### Generalized radial flow model (Barker, 1988)

This flow model handles flow dimensions between 1 (linear) and 3 (spherical). For the Theis solution a flow dimension of 2 (radial) can be used.  In addition, non-integral flow dimensions can also be used. The goal of these functions are to function with large datasets and calibration techniques.

### Parameters

```{r grfparameters, echo = TRUE}
times <- 1:2000
flow_rates <- c(abs(rnorm(1000)) / 1000, rep(0.0, 1000))

dat <- data.frame(times, flow_rates)

```

## Regularly spaced time series

The following step is designed for regularly spaced times series for both the flow rates and drawdown output times.

```{r grf, echo=TRUE}
# radial (flow_dimension = 2 Theis)
dd_rad <- recipe(times~flow_rates, dat) |>
  step_aquifer_grf(time = times, 
                   flow_rate = flow_rates, 
                   thickness = 1.0,
                   radius = 20, 
                   specific_storage = 1e-5,
                   hydraulic_conductivity = 1e-3,
                   flow_dimension = 2) |> 
  plate()


# linear (flow_dimension = 1)

dd_lin <- recipe(times~flow_rates, dat) |>
  step_aquifer_grf(time = times, 
                   flow_rate = flow_rates, 
                   thickness = 1.0,
                   radius = 20, 
                   specific_storage = 1e-5,
                   hydraulic_conductivity = 1e-3,
                   flow_dimension = 1) |> 
  plate()

# spherical (flow_dimension = 3)
dd_sph <- recipe(times~flow_rates, dat) |>
  step_aquifer_grf(time = times, 
                   flow_rate = flow_rates, 
                   thickness = 1.0,
                   radius = 20, 
                   specific_storage = 1e-5,
                   hydraulic_conductivity = 1e-3,
                   flow_dimension = 3) |> 
  plate()


```


```{r grfplot, echo=FALSE}

grf <- data.frame(times = times, 
                  flow_rates = flow_rates,
                  drawdown_rad = dd_rad$aquifer_grf,
                  drawdown_lin = dd_lin$aquifer_grf,
                  drawdown_sph = dd_sph$aquifer_grf
)


p1 <- plot_ly(grf, x = ~times, y = ~drawdown_lin,
        type = "scatter", mode = "lines", 
        yaxis = "y1", name = "linear")

p2 <- plot_ly(grf, x = ~times, y = ~drawdown_rad,
        type = "scatter", mode = "lines", 
        yaxis = "y1", name = "radial")

p3 <- plot_ly(grf, x = ~times, y = ~drawdown_sph,
        type = "scatter", mode = "lines", 
        yaxis = "y1", name = "spherical")

p4 <- plot_ly(grf, x = ~times, y = ~flow_rates,
        type = "scatter", mode = "lines", 
        yaxis = "y1", name = "flow rates")

subplot(p1, p2, p3, p4, shareX = TRUE, nrows = 4)


```
## References

Barker, J.A., 1988. A generalized radial flow model for hydraulic tests in fractured rock. Water Resour. Res. 24, 1796–1804.
